Library Learning in Neurally-Guided Bayesian Program Learning

### 2018-10-19 nips_2018.tex
tweaks to spacing

\author{
  Kevin Ellis\\MIT\\\texttt{ellisk@mit.edu}\And Lucas Morales\\MIT\\\texttt{lucas@lucasem.com}\And Mathias Sabl\'e-Meyer\\ENS Paris-Saclay\\\texttt{mathsm@mit.edu}
  \AND
  Armando Solar-Lezama\\MIT\\\texttt{asolar@csail.mit.edu}\And Joshua B. Tenenbaum\\MIT\\\texttt{jbt@mit.edu}
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

@@

Our algorithm takes inspiration from several ways that skilled human
programmers have learned to code: Skilled coders build libraries of
reusable subroutines that are shared across related programming tasks,
and can be composed to generate increasingly complex and powerful
subroutines.  In text editing, a good library should support routines
for splitting on characters, but also specialize these routines to
split on particular characters such as spaces or commas that are
frequently used to delimit substrings across tasks.  Skilled coders
also learn to recognize what kinds of programming idioms and library
routines would be useful for solving the task at hand, even if they
cannot instantly work out the details.  In text editing, one might
learn that if outputs are consistently shorter than inputs, removing
characters is likely to be part of the solution; if every output
contains a constant substring (e.g., ``Dr.''), inserting or appending
that constant string is likely to be a subroutine.

@@

How should we get the data to train $q$?
This is nonobvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from (program, task) pairs).
One approach is to sample programs from the DSL,
run them to get their input/outputs,
and then train $q$ to predict the program from the input/outputs.
This is like how a Helmholtz machine
trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
The advantage of ``Helmholtz machine'' training is that
we can draw unlimited samples from the DSL,
training on a large amount of data.
Another approach is
 self-supervised learning,
training $q$ on the (program, task)
pairs discovered by search.
The advantage of self-supervised learning
is that the training data is much higher quality,
because we are training on the actual tasks.
Due to these complementary advantages,
we train on both these sources of data.

@@

\begin{wrapfigure}{r}{0.55\textwidth}\centering
  \begin{tabular}{lll}
    \toprule
    Name & Input & Output \\\midrule
    repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
    keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
    product & [7\, 1\, 6\, 2] & 84 \\
    \bottomrule
  \end{tabular}
  \captionof{table}{Some tasks in our list function domain. See the
  supplement for the complete data set.}\label{listexamples}
\end{wrapfigure}
We created 236 human-interpretable list manipulation tasks, each with 15
input/output examples (Tbl.~\ref{listexamples}).
Our data set is interesting in three major ways: many of the tasks require
complex solutions; the tasks were not generated from some latent DSL, and
the agent must learn to solve these complicated problems from only 236
tasks.
Our data set assumes arithmetic operations as well as sequence operations,
so we additionally provide our system with the following arithmetic
primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
\code{is-prime}.

@@

We compare with ablations of our model on held out tasks.
The purpose of this ablation study is 
both to examine the role of each component of \systemEnding,
as well as to compare with
prior approaches in the literature:
a head-to-head
comparison of program synthesizers is complicated by the fact that
each system, including ours, makes idiosyncratic 
assumptions about the space of programs and the statement of tasks.
\begin{wraptable}{r}{0.55\textwidth}
\vspace{-5pt}
%% This radically shrinks the table
\tabcolsep=2.5pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lcccccc}
  \toprule& Ours& 
    No NN & SE&NPS & PCFG & Enum \\
  \midrule
  \multicolumn{7}{c}{\emph{List Processing}}\\
  \midrule
  \% solved&\textbf{94\%} & 
  79\% &71\%&35\%&62\%&37\%\\
  Solve time&  88s&39s&11s&35s&44s&20s\\
  \midrule
  \multicolumn{7}{c}{\emph{Text Editing}}\\
  \midrule
  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
  Solve time& 29s&65s&38s&80s&--&235s\\
  \midrule
  \multicolumn{7}{c}{\emph{Symbolic Regression}}\\
  \midrule
  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
  Solve time&  24s& 40s  &28s&31s&55s&29s\\
  \bottomrule
  \end{tabular}
\caption{\% held-out test tasks solved. Solve time: averaged over solved
  tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
  %structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
Nevertheless, much prior work can be modeled within our setup. 
We compare with the following ablations (Tbl~\ref{baselineComparisons};
Fig~\ref{learningCurves}):
\\\noindent \textbf{No NN:} lesions the recognition model.
\\\noindent \textbf{NPS}, which does not learn the DSL,
instead learning the recognition model
from samples drawn from the fixed DSL.
We call this NPS (Neural Program Synthesis)
because this is closest to how
RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL. This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first exploration step.

%For each domain,
We are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
Tbl.~\ref{baselineComparisons}
compares our model against these alternatives.
We consistently
improve on the baselines,
and also find that lesioning the recognition model
impairs the convergence of the algorithm,
causing it to hit a `plateau' after which it stops solving new tasks after an initial spurt of learning (Fig.~\ref{learningCurves}) -- without the neural network, search becomes intractable.
This `plateau' supports a view of the recognition model as a way of amortizing the cost of search.

>>>>

\author{
  Kevin Ellis\\MIT\\\texttt{ellisk@mit.edu}\And Lucas Morales\\MIT\\\texttt{lucasem@mit.edu}\And Mathias Sabl\'e-Meyer\\ENS Paris-Saclay\\\texttt{mathsm@mit.edu}
  \AND
  Armando Solar-Lezama\\MIT\\\texttt{asolar@csail.mit.edu}\And Joshua B. Tenenbaum\\MIT\\\texttt{jbt@mit.edu}
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

@@

Our algorithm takes inspiration from several ways that skilled human
programmers have learned to code: skilled coders build libraries of
reusable subroutines that are shared across related programming tasks,
and can be composed to generate increasingly complex and powerful
subroutines.  In text editing, a good library should support routines
for splitting on characters, but also specialize these routines to
split on particular characters such as spaces or commas that are
frequently used to delimit substrings across tasks.  Skilled coders
also learn to recognize what kinds of programming idioms and library
routines would be useful for solving the task at hand, even if they
cannot instantly work out the details.  In text editing, one might
learn that if outputs are consistently shorter than inputs, removing
characters is likely to be part of the solution; if every output
contains a constant substring (e.g., ``Dr.''), inserting or appending
that constant string is likely to be a subroutine.

@@

How should we get the data to train $q$?
This is non-obvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from task/program pairs).
One approach is to sample programs from the DSL,
run them to get their input/outputs,
and then train $q$ to predict the program from the input/outputs.
This is like how a Helmholtz machine
trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
The advantage of ``Helmholtz machine'' training is that
we can draw unlimited samples from the DSL,
training on a large amount of data.
Another approach is
 self-supervised learning,
training $q$ on the (program, task)
pairs discovered by search.
The advantage of self-supervised learning
is that the training data is much higher quality,
because we are training on the actual tasks.
Due to these complementary advantages,
we train on both these sources of data.

@@

\begin{wrapfigure}{r}{0.55\textwidth}\centering
  \begin{tabular}{lll}
    \toprule
    Name & Input & Output \\\midrule
    repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
    keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
    product & [7\, 1\, 6\, 2] & 84 \\
    \bottomrule
  \end{tabular}
  \captionof{table}{Some tasks in our list function domain. See the
  supplement for the complete data set.}\label{listexamples}
\end{wrapfigure}
We created 236 human-interpretable list manipulation tasks, each with 15
input/output examples (Tbl.~\ref{listexamples}).
Our data set is interesting in three major ways: many of the tasks require
complex solutions; the tasks were not generated from some latent DSL; and
the agent must learn to solve these complicated problems from only 236
tasks.
Our data set assumes arithmetic operations as well as sequence operations,
so we additionally provide our system with the following arithmetic
primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
\code{is-prime}.

@@

We compare with ablations of our model on held out tasks.
The purpose of this ablation study is 
both to examine the role of each component of \systemEnding,
as well as to compare with
prior approaches in the literature:
a head-to-head
comparison of program synthesizers is complicated by the fact that
each system, including ours, makes idiosyncratic 
assumptions about the space of programs and the statement of tasks.
\begin{wraptable}{r}{0.55\textwidth}
\vspace{8pt}
%% This radically shrinks the table
\tabcolsep=2.5pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lcccccc}
  \toprule& Ours& 
    No NN & SE&NPS & PCFG & Enum \\
  \midrule
  \multicolumn{7}{c}{\emph{List Processing}}\\
  \midrule
  \% solved&\textbf{94\%} & 
  79\% &71\%&35\%&62\%&37\%\\
  Solve time&  88s&39s&11s&35s&44s&20s\\
  \midrule
  \multicolumn{7}{c}{\emph{Text Editing}}\\
  \midrule
  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
  Solve time& 29s&49s % FIXME: 65 in the table but 49 on the graph? 65s
  &38s&80s&--&235s\\
  \midrule
  \multicolumn{7}{c}{\emph{Symbolic Regression}}\\
  \midrule
  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
  Solve time&  24s& 40s  &28s&31s&55s&29s\\
  \bottomrule
  \end{tabular}
\caption{\% held-out test tasks solved. Solve time: averaged over solved
  tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
  %structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0cm}\label{baselineComparisons} \end{wraptable}
Nevertheless, much prior work can be modeled within our setup. 
We compare with the following ablations (Tbl~\ref{baselineComparisons};
Fig~\ref{learningCurves}):
\\\noindent \textbf{No NN:} lesions the recognition model.
\\\noindent \textbf{NPS}, which does not learn the DSL,
instead learning the recognition model
from samples drawn from the fixed DSL.
We call this NPS (Neural Program Synthesis)
because this is closest to how
RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL. This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first exploration step.

%For each domain,
We are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
Tbl.~\ref{baselineComparisons}
compares our model against these alternatives.
We consistently
improve on the baselines,
and also find that lesioning the recognition model
impairs the convergence of the algorithm,
causing it to hit a lower `plateau' after which it stops solving new tasks, following an initial spurt of learning (Fig.~\ref{learningCurves}) -- without the neural network, search becomes intractable.
This lowered `plateau' supports a view of the recognition model as a way of amortizing the cost of search.
